% !TeX spellcheck = en_GB 

\section{\label{sec::setupSystem}Setting up the System}
In this section we find a relaxation of the problem stated in section \ref{sec::problem}. Multiplying both sides of (\ref{eqn::strongForm}) with an arbitrary test function $v\in V:=H^1(\Omega)$ and integrating over the domain, we get the weak formulation:
\begin{equation*}
        \int\limits_{\Omega} \frac{\partial u}{\partial t} v = \int\limits_{\Omega} (\nabla(\alpha\nabla u)) v.
\end{equation*}
Integrating the right hand side by parts will result in:
\begin{equation*}
        \int\limits_{\Omega} \frac{\partial u}{\partial t} v = -\int\limits_{\Omega} \alpha(\nabla u) \cdot (\nabla v) + \int\limits_{\partial \Omega} \alpha v (\nabla u) \cdot \nu.
\end{equation*}
Using
\begin{equation*}
        (\nabla u)\cdot \nu = \frac{\partial u}{\partial \nu} = g
\end{equation*}
from (\ref{eqn::neumann}) leads to:
\begin{equation}
        \label{eqn::weakForm}
        \int\limits_{\Omega} \frac{\partial u}{\partial t} v = -\int\limits_{\Omega} \alpha(\nabla u) \cdot (\nabla v) +  \int\limits_{\partial\Omega^N} \alpha gv + \int\limits_{\partial \Omega^D} \alpha v (\nabla u) \cdot \nu.
\end{equation}
Using the above we can now define the bilinear form $a$ and the linear operator $L$:
\begin{align*}
        a(u,v)&=\int\limits_{\Omega} \left(\frac{\partial u}{\partial t} v +  \alpha(\nabla u) \cdot (\nabla v) \right) \\
        L(v) &= \int\limits_{\partial\Omega^N} \alpha gv.
\end{align*}
From now on we assume that $a$ is bounded and coercive and that $L$ is bounded. This will assure that we can apply Lax-Milgram \cite{quarteroni2009numerical}, which implies the heat equation has a unique solution.

To solve the weak formulation (\ref{eqn::weakForm}) numerically we discretise our domain $\Omega$. Our notation will follow \cite{quarteroni2009numerical}. Also we will not state every step in detail, if you wish, to get deeper insights in the theory behind it, we also recommend reading \cite{quarteroni2009numerical}.

Let $\mathcal{T}_h$ be a set of non overlapping tetrahedrons covering $\Omega$ and $\mathcal{N}={N_1,\dots N_{n_h}}$ the nodes of this mesh. As the theory about this is not new, and not closely related, to our problem, we don't want to go into detail about this. The approximated domain is then $\Omega_h:=\bigcup_{K\in\mathcal{T}_h}K$.

As an approximation of the functions in $V$, we use the Galerkin method, searching only for functions in $X_h:=\{v_h\in C^0(\overline{\Omega}_h) : v_h|_K \text{ linear } \forall K\in \mathcal{T}_h\}\subset V$. These are the continuous functions on $\Omega_h$, that are piecewise linear on each tetrahedron.

A basis for this space is given by the characteristic Lagrangian functions $\phi_j\in\ X_h, j=1,\dots n_h$, with $\phi_j(N_i) = \delta_{ij}$. So we can write every $v_h\in X_h$ in the following way:
\begin{equation*}
        v_h(x) = \sum_{j=1}^{j=n_h}v_j\phi_j.
\end{equation*}

On this space the equation (\ref{eqn::weakForm}) is equivalent to the following:
\begin{equation}
\label{eqn::discreticedWeakForm}
\begin{aligned}
        \int\limits_{\Omega} \sum_i \frac{\partial u_i}{\partial t} \phi_i \phi_j = &-\int\limits_{\Omega} \sum_i \alpha u_i(\nabla \phi_i) \cdot (\nabla \phi_j) \\
        &+ \int\limits_{\partial\Omega^N} \alpha g\phi_j \\
        &+ \int\limits_{\partial\Omega^D} \sum_i \alpha u_i (\nabla \phi_i) \cdot \nu \phi_j & \forall j.
\end{aligned}
\end{equation}

Considering the case of homogeneous Dirichlet conditions (i.e. $u^D=0$), we only have to search on the subspace $V_h:=\mathring{X}_h:=\{v_h\in X_h : v_h|_{\partial\Omega_h^D} = 0\}$. Let WLOG be the last indices $n_D,\dots n_h$, the indices of the nodes on the Dirichlet boundary. As $v_h\in V_h$ leads to $v_j = 0, \forall j\geq n_D$, (\ref{eqn::discreticedWeakForm}) becomes:
\begin{equation}
\label{eqn::homogeneousForm}
\begin{aligned}
        \int\limits_{\Omega} \sum_i \frac{\partial u_i}{\partial t} \phi_i \phi_j = &-\int\limits_{\Omega} \sum_i \alpha u_i(\nabla \phi_i) \cdot (\nabla \phi_j) \\
        &+ \int\limits_{\partial\Omega^N} \alpha g\phi_j
         & \forall j < n_D.
\end{aligned}
\end{equation}

We can reduce the non-homogeneous case to the homogeneous one, by introducing a lifting $R_g\in X_h$ as follows:
\begin{equation*}
        R_g(x) := \sum_{i=n_D}^{n_h}d_i \phi_i(x),
\end{equation*}
where $d_i:=u^D(N_i)$. With the homogeneous solution
\begin{equation*}
        \mathring{u} := \sum_{i=1}^{n_D-1}u_i \phi_i(x),
\end{equation*}
the final solution is given by
\begin{equation}
        \label{eqn::fullu}
        u = \mathring{u} + R_g.
\end{equation}


To find this homogeneous solution, we insert (\ref{eqn::fullu}) in (\ref{eqn::homogeneousForm}) and get:
\begin{equation}
\label{eqn::finalDiscreticedForm}
\begin{aligned}
        \int\limits_{\Omega} \sum_{i=1}^{n_D-1} \frac{\partial u_i}{\partial t} \phi_i \phi_j = &-\int\limits_{\Omega} \sum_{i=1}^{n_D-1} \alpha u_i(\nabla \phi_i) \cdot (\nabla \phi_j) \\
        &+ \int\limits_{\partial\Omega^N} \alpha g\phi_j \\
        &-        \int\limits_{\Omega} \sum_{i=n_D}^{n_h}\left( \dot{d}_i \phi_i \phi_j  + \alpha d_i(\nabla \phi_i) \cdot (\nabla \phi_j)\right)
         & \forall j,
\end{aligned}
\end{equation}
with $\dot{d}_i:=\frac{\partial u^D}{\partial t}(N_i)$.

We now define the matrices $M$, $A$ and the vectors $N$, $D$ by:
\begin{align*}
        M_{ij} &:=\int\limits_{\Omega} \phi_i \phi_j \\
        A_{ij} &:=\int\limits_{\Omega} \alpha (\nabla \phi_i) \cdot (\nabla \phi_j) \\
        N_{j} &:=\int\limits_{\partial\Omega^N} \alpha g\phi_j \\
        D_{j} &:=\int\limits_{\Omega} \sum_{k=n_D}^{n_h}\left( \dot{d}_k \phi_k \phi_j  + \alpha d_k(\nabla \phi_k) \cdot (\nabla \phi_j)\right),
\end{align*}
where $i,j=1,\dots (n_D-1)$. Using these we get the system of ordinary differential equations:
\begin{equation}
        \label{eqn::matrixForm}
        M\frac{\partial u}{\partial t} = -A u + N - D.
\end{equation}
How we solve these will be considered in the following section.

For the practical part we implemented functions, which can calculate these matrices and vectors. To do so we use the transformation properties of the barycentric coordinates from the basic element (triangle or higher dimensional equivalent). As we wanted to be able to solve the problem as general as possible, we allow $g$ and $d$ to be space and time dependent. If one of them depends on time, the corresponding vectors has to be calculated every time step in the solving progress of the differential equation (section \ref{sec::odesolver}). The thermal diffusivity $\alpha$ can differ between the elements, but we assume it to be constant within each.
